Copulas and their potential for ecology

### 2019-12-16 AER_Responses.tex
working on referee comments, second referee now included

>>>>

Referee 2

Review of Ghosh et al.: Copulas and the Potential for Ecology
I would like to thank the authors for their manuscript entitled "Copulas and their potential for ecology".  The manuscript provides an introduction to the concept of copulas and their utility for application in studying the natures of variable dependency in ecology.  The authors demonstrate that the most commonly used models of inference, correlation and regression, only encapsulate certain features of the dependency between variables and the author introduce a more complete method of inference to an ecologist audience.
I must first state that I must give fair warning that, whilst I have heard about copulas before, I have never applied them to my own research and I certainly cannot claim to be an expert in this area.  I am therefore probably the target audience of this paper.  This means that whilst I can give constructive feedback on the author's introduction of the concepts to an ecologist readership and can follow the mathematical background that they have provided, I cannot necessarily state definitively whether the authors have reproduced a fair portrayal of the theory of copulas in this manuscript.
However, despite not being a specialist on the subject, I found the presentation of the manuscript reasonably easy to follow and the authors should be commended for making what is quite a dense subject approachable.  Each theoretical claim is backed by well-documented supplementary materials which allows for digestion of the key concepts without being side-tracked by dense mathematical derivations.

>>>Thank you for the positive feedback.

The authors detail a large number of example ecological applications and detail a strong case for the use of copula families with left- or right- tail associations.  This is a very good demonstration where the use of copulas can benefit the ecological investigator far above the traditional methods.

>>>Thank you for the positive feedback.

I would however appreciate a discussion of how copula-based methods compare with regression models with a heteroscedastic error term.  It seems that these models would also be able to cope with left- or right- tail associations in a bivariate application.  Given the rise of hierarchical Bayesian models in ecology, before reading this paper, my (maybe somewhat naiive) approach would be to apply a standard regression model but allow for the variance in the error distribution to also be determined by the predictor variable.

>>><DAN: to add some discussion material>

In my view this paper represents an impressive contribution to statistical ecology and I am very happy to be introduced to a new and exciting avenue of analysis.  Indeed, the thing I hate most about the manuscript is that it has spurred me on to spend hundreds of dollars on new academic books on copula modelling.

>>>Thank you for the positive feedback. This was exactly our goal!

What follows is a small set of very minor points that are all highly subjective and whether or not they are addressed does not, in my view, preclude the manuscript of being published:
o    The authors use the term "distribution function" to represent "cumulative distribution function".  My feeling is that ecologists in general are more familiar with the term "cumulative distribution function" (which thereafter can be abbreviated to "CDF") as the former is more likely to be mistaken for the density function.

>>><DAN to do>

o    The manuscript could benefit from a little more detail in the introduction as to how this material goes further than that in other ecological introductions to copulas.  For example, Anderson et al. 2018 also contains an introduction to the theory of copulas so if might be worth explicitly pointing out how this paper goes further in explaining and introducing the concept to an ecologist audience.

>>><DAN to do>

o    A table giving a run-down of the notation used in the manuscript might be useful for the more mathematically-adverse reader.  A sort of notation cheat-sheet for easy notation lookup would be useful for readers so that they can concentrate on the new concepts rather than looking up notation definitions.

o    The authors make reference to many families of copula models (a total of 16 families) and, whilst there is a graphical illustration of the shapes of the copulas, we don't see the functional forms of these copulas.  A table either in the main text or supplementary materials giving these would be appreciated.

### 2019-12-16 Paper.Rmd
working on referee comments, second referee now included

<!--***DAN: remianing to do:
DONE 1) Edit abstract
DONE 2) Go through the whole thing one more time
DONE 3) send it out to coauthors (done) and maybe get feedback
DONE 3.5) ESP FROM lAWRENCE
DONE 3.6) Incorporate Shya feedback
DONE 3.7) Incorporate Reid feedback
DONE 4)make sure to say something about tropical systems
DONE 5) maybe write to the editors saying this is coming soon?
DONE 6) In the Discussion, you maybe need to point out not only that we
use only bivariate copulas (and there are multivariate ones), but also
that we use only continuous marginals. Probably an adequate approximation
for our data, but see the other papers I've seen recently which use count data,
and discrete marginals. Approaches such as that will be necessary for some 
datasets.
DONE 7) change the box around the Cedar Creek data in the main fig
DONE 8) say in Discussion that you could have used non-ranks for the nonparam stats - this
is the stuff the Ecol Mongr referee talked about. Maybe this should go in Background instead?
DONE 9) maybe go back and read the Ecol Monogr reviews again
DONE 10) try to emphasize early on that the focus is on positively associated variables, so 'lower tail' means the lower tail of both variables and it is not necessary to worry about negative relationships (in practice we would always transform these for ease).   The dfinitions of tail dependence we use cannot usefully be applied to untransformed data in fig 1e,f.
DONE 11) Number the sections and subsections in the main text and change all the section 
references to use those numbers. Search for "section" (DONE) also for "methods" (DONE)
-->

@@

We introduce some standard copula families, as examples of copulas and also
because we will fit these families to data later. The bivariate normal
copula family, which consists of copulas of bivariate normal
distributions, is a one-parameter family. The parameter, which we here call $p$,
corresponds to the correlation
of the related bivariate normal distribution, and controls the degree of association
between the variables. A normal copula with $p=0.7$ was already introduced (Fig.
\ref{PedagogFigCopThry}A). The formula for the
copula is $\Phi_{2,p}(\Phi_{1}^{-1}(u),\Phi_{1}^{-1}(v))$, where $\Phi_1^{-1}$ is the inverse of the
cummulative dstribution function (cdf) of a (univariate) standard 
normal and $\Phi_{2,p}$ is the joint cdf of a
bivariate normal distribution with mean $(0,0)$ and covariance matrix having $1$s on the diagonal
and $p$ in the off diagonal positions. However, formulas for copulas are often not as 
directly informative about copula properties as probability density function (pdf) 
pictures; Fig. \ref{figped_N} has 
example pdfs corresponding to bivariate normal copulas for various values of $p$.
Note that the pdfs are symmetric
around the diagonal line $v=-u+1$ (Fig. \ref{figped_N}), so normal copulas have symmetric 
associations between the two variables in the left and right tails. 
The Clayton copula family,
of which we have already pictured an example (Fig.
\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to normal copulas, Clayton copulas
have stronger left- than right-tail association. The formula is 
$\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 
less direct intuition than do example pdfs (Fig. \ref{figped_C}). Higher values of the parameter,
$p$ produce copulas from the Clayton family with higher Kendall or Spearman correlation. The survival
Clayton family is the symmetric opposite of the Clayton family, showing
stronger right- than left tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
family is a two-parameter family, thereby providing more flexibility: for some parameters
BB1 copulas have stronger left- than right-tail association, and for others the reverse
(Fig. \ref{figped_BB1}). See @nelsen2006_copula and @joe2014_dependence for further
information on these and other copula families.

>>>>

<!--***DAN: remianing to do:
-->

@@

We introduce some standard copula families, as examples of copulas and also
because we will fit these families to data later. The bivariate normal
copula family, which consists of copulas of bivariate normal
distributions, is a one-parameter family. The parameter, which we here call $p$,
corresponds to the correlation
of the related bivariate normal distribution, and controls the degree of association
between the variables. A normal copula with $p=0.7$ was already introduced (Fig.
\ref{PedagogFigCopThry}A). The formula for the
copula is $\Phi_{2,p}(\Phi_{1}^{-1}(u),\Phi_{1}^{-1}(v))$, where $\Phi_1^{-1}$ is the inverse of the
cummulative dstribution function (cdf) of a (univariate) standard 
normal and $\Phi_{2,p}$ is the joint cdf of a
bivariate normal distribution with mean $(0,0)$ and covariance matrix having $1$s on the diagonal
and $p$ in the off diagonal positions. However, formulas for copulas are often not as 
directly informative about copula properties as probability density function (pdf) 
pictures; Fig. \ref{figped_N} has 
example pdfs corresponding to bivariate normal copulas for various values of $p$.
Note that the pdfs are symmetric
around the diagonal line $v=-u+1$ (Fig. \ref{figped_N}), so normal copulas have symmetric 
associations between the two variables in the left and right tails. 
The Clayton copula family,
of which we have already pictured an example (Fig.
\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to normal copulas, Clayton copulas
have stronger left- than right-tail association. The formula is 
<!--***DAN: is it clear implicitly from what came before that this is the cdf or what?-->
$\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 
less direct intuition than do example pdfs (Fig. \ref{figped_C}). Higher values of the parameter,
$p$ produce copulas from the Clayton family with higher Kendall or Spearman correlation. The survival
Clayton family is the symmetric opposite of the Clayton family, showing
stronger right- than left tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
family is a two-parameter family, thereby providing more flexibility: for some parameters
BB1 copulas have stronger left- than right-tail association, and for others the reverse
(Fig. \ref{figped_BB1}). See @nelsen2006_copula and @joe2014_dependence for further
information on these and other copula families.
